---
title: "Title Goes Here"
author:
  - Jane Doe:
      correspondence: "yes"
      email: janedoe@fosg.org
      orcid: 0000-0003-1689-0557
      institute:
        - fosg
        - fop
  - John Q. Doe:
      institute: fosg
      orcid: 0000-0003-1689-0558
  - Peder Ås:
      institute: fosg
      orcid: 0000-0003-1689-0559
  - Juan Pérez:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Acme Corporation
  - Max Mustermann:
      orcid: 0000-0003-1689-0552
institute:
  - fosg:
      name: Formatting Open Science Group
      address: 23 Science Street, Eureka, Mississippi, USA
  - fop: Federation of Planets
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The actual document text starts here: -->

# Introduction

Here is a citation [@Marwick2017] Here is another citation @brumm2018

Even more citations [@marwick2016].

<!-- visual mode for citations + must have empty spaces when making those citations / @ + van den burgh showing up as bergh -->

# Background

On the Wallacean island of Sulawesi, Indonesia, there have been various cave sites and rockshelters excavated across the island with evidence of late Pleistocene to early Holocene lithic technology. Deep excavations performed at the site of Talepu yielded 270 lithic artefacts from trench two and 41 from trench four, including four in situ lithics. Multiple-elevated-temperature post-infrared infrared stimulated luminescence (MET-pIRIR) ages of sediments indicate dates for lithic artefacts at Talepu between 118kyr to 194 kyr, making the site of Talepu one of the oldest lithic sites on Sulawesi. The predominant source of raw material for these lithics was silicified limestone cobbles. Approximately 13% of the assemblage consisted of cores, reduced both unifacially (42%) and bifacially (58%), and flakes that were similarly reduced unifacially (60%) and bifacially (40%) [@vandenbergh2016]. The authors note that it is unlikely that there was a specific form in mind when the cobbles were being shaped into tools and that the flakes were likely an intentional byproduct to source more flakes, but no further analysis was made on any changes in the lithic technology over time.

Later in the Pleistocene, at around 50 ka BP, lithics appear at Leang Bulu Bettue in the southwestern peninsula of Sulawesi. Excavations conducted by @brumm2017 yielded two distinct phases spanning from the Pleistocene period, (∼50−16 ka cal BP), to the Holocene (\~1.7-4.8 ka cal BP). Phase I, spanning layers 4a-5 and correlating to the Pleistocene, was dated using several different methods: AMS C14-dating of freshwater Mollusca shells, laser ablation U-series dating of a bovid molar, U-series isotope analysis of stalagmites, and optical dating of feldspar grains [@brumm2017; @li2016]. The oldest of the layers in Phase I, layer 5, is where the oldest lithic artefacts were found. Made of limestone and volcanic material, these are associated with pIRIR and U-series ages of 50-40 ka. Phase II, encompassing the uppermost layers, 1-3 and correlating with the Holocene, was dated mostly using AMS C14-dating of both freshwater shells and charcoal. The recency of the Phase 2 deposit is further reinforced by the presence of Neolithic and Historic-era pottery and lithics [@perston2022].

The lithic assemblages excavated at Leang Bulu Bettue reveal notable changes in technology and typology over time. In Layer 5 (\~50 ka) of Phase I, the artefacts predominantly consisted of flakes measuring \~50-55 mm long [@perston2022]. There was a preference for locally available limestone for core production, with chert being less prevalent. Core reduction was minimal, but primarily involved freehand hard-hammer percussion, resulting in single-platform and multiplatform cores. A technological shift took place between the transition from the Pleistocene layer (layer 5) and the Holocene layer (4) of Phase I. This is a shift to the predominant use of chert for tools, along with more technical reduction methods, such as unifacial, bifacial, and bipolar flaking. Further, we see a shift in flake sizes, with the longest measurements being \~105-108 mm, nearly double the length of the layer 5 lithics. Unfortunately, the sample size and details provided for Phase II of the assemblage, encompassing the uppermost layers, are too small to fairly compare any technological changes. However, it is evident that the transition to layer 4 brought about a broader utilization of materials and increased complexity in flaking and reduction techniques.

Leang Burung 2, approximately 20 km north of Leang Bulu Bettue, provides a complex chronology of the technological evolution of lithic artefacts in the region. Stratigraphic dating of Layers A and I using Post-infrared infrared stimulated luminescence (pIRIR) of feldspars, U-series dating of a straw stalactite, and C14-dating of rear wall breccias yielded minimum ages between \~50-35 ka cal BP [@brumm2018]. While Layer A is characterized by the use of local limestone cobbles and hard hammer percussion, over half of the Layer I assemblage is made up of chert, followed by limestone and quartz [@glover1981]. The subsequent layers (V, IV, II, and III/IIIa-b) further evidence the transition to a predominantly chert-based technology and a continuation of hard-hammer reduction, in addition to, and less commonly, bipolar reduction [@brumm2018].

Leang Sakapao 1, another nearby site, presents a similar case in materials and technology. C14-dating of marine shells found at the lowest spits of the excavation (spits 5/6) yielded calibrated dates of \~30-29 ka cal BP [@bulbeck2004; @oconnor2014], placing the occupation of this site around a similar time to that of Leang Burung 2. Another similarity in technology which we see at this site is the primary use of chert, along with hard-hammer reduction and the occasional use of bipolar reduction (n=5) [@perston2022].

Beyond a shift from the primary usage of limestone cobbles to chert (layer A), minimal technological change appears to have occurred at Leang Burung 2 and Leang Sakapao 1 over time. However, it should be noted that the small sample size and minimal dating of samples at Leang Sakapao 1 makes it difficult to discern any major technological changes that may have occurred over time. Further, the site is considered to have poor accessibility, due to its high elevation, in comparison to Leang Burung 2, which may account for the limitation of material to being predominantly chert throughout the entire assemblage. We see further evidence of how cave accessibility may have impacted the makeup of the assemblages through the wider range of flakes (Levallois points, a retouched point, and retouched scrapers) present at Leang Burung 2 ([@bulbeck2004; @glover1981; @oconnor2014].

```{=html}
<!--
last background paragraph notes

-should i make note of flakes with silica gloss ? didn't really appear to be a major detail
-trouble importing Glover 1981 citation, just added so maybe try again later
-(n=(number)) proper in research papers ?
-expansions needed in final paragraph?
-->
```
# Methods

# Results

<!-- Here's some example analysis code: -->

```{r}
#| label: get-data
#| eval: false
# Note the path that we need to use to access our data files when rendering this document
my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
```

```{r}
#| label: fig-demo-plot
#| fig-cap: "A plot of random numbers"
plot(rnorm(10))
```

@fig-demo-plot shows how we can have a caption and cross-reference for a plot. Note that figure label and cross-references must both be prefixed with `fig-`

```{r}
#| label: demo-inline-code
#| echo: false
x <- round(pi, 2)
```

Here is an example of inline code `r x` in the middle of a sentence.

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
#| label: colophon
#| cache: false

# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```
