---
title: "Title Goes Here"
author:
  - Jane Doe:
      correspondence: "yes"
      email: janedoe@fosg.org
      orcid: 0000-0003-1689-0557
      institute:
        - fosg
        - fop
  - John Q. Doe:
      institute: fosg
      orcid: 0000-0003-1689-0558
  - Peder Ås:
      institute: fosg
      orcid: 0000-0003-1689-0559
  - Juan Pérez:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Acme Corporation
  - Max Mustermann:
      orcid: 0000-0003-1689-0552
institute:
  - fosg:
      name: Formatting Open Science Group
      address: 23 Science Street, Eureka, Mississippi, USA
  - fop: Federation of Planets
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The actual document text starts here: -->

# Introduction

Here is a citation [@Marwick2017] Here is another citation @brumm2018

Even more citations [@marwick2016].

# Background

On the Wallacean island of Sulawesi, Indonesia, there have been various cave sites and rockshelters excavated across the island with evidence of late Pleistocene to early Holocene lithic technology. Deep excavations performed at the site of Talepu yielded 270 lithic artefacts from trench two and 41 from trench four, including four in situ lithics. Multiple-elevated-temperature post-infrared infrared stimulated luminescence (MET-pIRIR) ages of sediments indicate dates for lithic artefacts at Talepu between 118kyr to 194 kyr, making the site of Talepu one of the oldest lithic sites on Sulawesi. The predominant source of raw material for these lithics was silicified limestone cobbles. Approximately 13% of the assemblage consisted of cores, reduced both unifacially (42%) and bifacially (58%), and flakes that were similarly reduced unifacially (60%) and bifacially (40%) (van den Bergh et al., 2016). The authors note that it is unlikely that there was a specific form in mind when the cobbles were being shaped into tools and that the flakes were likely an intentional byproduct to source more flakes, but no further analysis was made on any changes in the lithic technology over time.  

Later in the Pleistocene, at around 50 ka BP, lithics appear at Leang Bulu Bettue in the southwestern peninsula of Sulawesi. Excavations conducted by Brumm, et al. (2017; Li et al., 2016) yielded two distinct phases spanning from the Pleistocene period, (∼50−16 ka cal BP), to the Holocene (~1.7-4.8 ka cal BP). Phase I, spanning layers 4a-5 and correlating to the Pleistocene, was dated using several different methods: AMS C14-dating of freshwater Mollusca shells, laser ablation U-series dating of a bovid molar, U-series isotope analysis of stalagmites, and optical dating of feldspar grains (Brumm et al., 2017; Li et al., 2016). The oldest of the layers in Phase I, layer 5, is where the oldest lithic artefacts were found. Made of limestone and volcanic material, these are associated with pIRIR and U-series ages of 50–40 ka. Phase II, encompassing the uppermost layers, 1-3 and correlating with the Holocene, was dated mostly using AMS C14-dating of both freshwater shells and charcoal. The recency of the Phase 2 deposit is further reinforced by the presence of Neolithic and Historic-era pottery and lithics (Perston et al., 2022).


# Methods

# Results

<!-- Here's some example analysis code: -->

```{r}
#| label: get-data
#| eval: false
# Note the path that we need to use to access our data files when rendering this document
my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
```

```{r}
#| label: fig-demo-plot
#| fig-cap: "A plot of random numbers"
plot(rnorm(10))
```

@fig-demo-plot shows how we can have a caption and cross-reference for a plot. Note that figure label and cross-references must both be prefixed with `fig-`

```{r}
#| label: demo-inline-code
#| echo: false
x <- round(pi, 2)
```

Here is an example of inline code `r x` in the middle of a sentence.

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
#| label: colophon
#| cache: false

# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```
